Network Origins of Social Distinction

### 2025-07-20 manuscript.tex
Updates from Overleaf

\subsection{Distinction Centrality's Measurement Parameters}
In Figure 1, we present a variety of toy models that compare the effects of network interconnections on the status and constraint of actors. In each model, there are seven nodes, and node 1 is the best-connected node in the network. Thus, node 1 always has an eigenvector centrality of 1. However, the distinction of node 1 changes based on the interconnection of the other nodes, demonstrating how distinction centrality captures the question of constraint exerted by neighbors.

@@

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/star.png}
            \caption{Star Graph}
            \label{fig:star}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
            \caption{Structural Fold}
            \label{fig:sf}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/intra.png}
            \caption{2-Connected Subgraphs}
            \label{fig:intra}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/inter.png}
            \caption{2-Connected Subgraphs with Long Tie}
            \label{fig:inter}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle.png}
            \caption{Circle Graph}
            \label{fig:circle}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle-plus-triangle.png}
            \caption{Circle Graph with Closure}
            \label{fig:circleplus}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle-plus-long.png}
            \caption{Circle Graph with Long Tie}
            \label{fig:circleplus}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/clique.png}
            \caption{Clique Graph}
            \label{fig:clique}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/kite.png}
            \caption{Kite Graph}
            \label{fig:kite}
    \end{subfigure}
    \caption{Distinction Centrality in toy graphs. Nodes with the highest distinction are shown in red, nodes with positive distinction are shown in blue, nodes with negative distinction are shown in tan, and nodes with distinction close to zero are shown in purple.}
    \label{fig:toys}
\end{figure}

@@

\subsection{Distinction with Structural Holes vs. Structural Folds}

%Clique addition experiment

%Add cliques to star graph and see what happens

%See what happens with the caveman

%Placehold graphs
%\begin{figure}
%   \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
%    \centering
%     \begin{subfigure}[b]{0.45\textwidth}
%        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
%            \caption{Star Graph}
%            \label{fig:Placehold}
%    \end{subfigure}
%     \begin{subfigure}[b]{0.45\textwidth}
%        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
%            \caption{Structural Fold}
%            \label{fig:Placehold}
%    \end{subfigure}
%    \caption{Structural Holes with more added subgraphs.}
%    \label{fig:toys}
%\end{figure}

\subsection{Long Ties and Hierarchy}
This section highlights how the hierarchy is really formed out of the redundant connections of the subgraphs. While a star graph clearly creates a hierarchy, when additional weight is added to the broker's alters, the hierarchy can be flipped, especially when they form denser subgraphs. This is in keeping with the observed structural holes above.

@@

%NOTES FROM 7|19|25 Meeting
%***Triadic closure as distinction maximization*** 
%Strogatz is random, preferential attachment of closure
%Status homophilous (enh) attachment vs preferential vs random 
%I wonder how much of balance theory is really just distinction maximization?
%Non psychological theory polarity

%3 versions
%-Psychology
%-Polarity, valenced ties 
%-Status distinction

%*Tie that cuts the longest shortest path Science, 
%https://www.science.org/doi/full/10.1126/science.aau9735
%Distinction homophily, chicken or egg of

>>>>

\subsection{Distinction Centrality's Measurement Parameters}
Below we present a variety of toy model to explore the relationship between network shape and density and the calculated measure of network distinction.

\subsubsection{Distinction and Brokerage: Structural Holes vs. Structural Folds}
In Figure 1, we present a variety of toy models that compare the effects of network interconnections on the status and constraint of actors. In each model, there are seven nodes, and node 1 is the best-connected node in the network. Thus, node 1 always has an eigenvector centrality of 1. However, the distinction of node 1 changes based on the interconnection of the other nodes, demonstrating how distinction centrality captures the question of constraint exerted by neighbors.

@@

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.35\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/star.png}
            \caption{Star Graph}
            \label{fig:star}
    \end{subfigure} \\
     \begin{subfigure}[b]{0.35\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/sf.png}
            \caption{Structural Fold}
            \label{fig:sf}
    \end{subfigure} 
    \begin{subfigure}[b]{0.35\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/intra.png}
            \caption{2-Connected Subgraphs}
            \label{fig:intra}
    \end{subfigure}
    \begin{subfigure}[b]{0.35\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/inter.png}
            \caption{2-Connected Subgraphs with Long Tie}
            \label{fig:inter}
    \end{subfigure}
    \begin{subfigure}[b]{0.35\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/wheel.png}
            \caption{Wheel Graph}
            \label{fig:wheel}
    \end{subfigure}

    \caption{Distinction Centrality in toy graphs. Nodes with the highest distinction are shown in red, nodes with positive distinction are shown in blue, nodes with negative distinction are shown in tan, and nodes with distinction close to zero are shown in purple.}
    \label{fig:toys}
\end{figure}

@@

\subsubsection{Circle Graphs and Triadic Closure Incentives}
Here be dragons.

%Strogatz is random, preferential attachment of closure
%Status homophilous (enh) attachment vs preferential vs random

%***Triadic closure as distinction maximization*** 
%Triadic closure 3 versions
%-Psychology
%-Balance version response triadic closure, Polarity, valenced ties  balance was the reason 
%-Status distinction 
%I wonder how much of balance theory is really just distinction maximization?
%Knights of the Round table, why equality conditions are fragile
%kite graph-outsider breaks mechanical solidarity
%mechanical solidarity is really an outcome of the network structure (Durkheim kind of %said that but Omar said it better). Try to get inequality in an organic solidarity system but that's unstable? (might have written that down wrong). 
%mechanical solidarity benefits from the stranger. more theoretical than empirical. 
%actually anthropologists ruined everything in the villages
%also somewhat the Marco Polo vibe, Shogun (the violation of markers of distinction, %didn't matter), The Last Samurai
%theory of subgraph cliques rather than full graph as clique
%*Tie that cuts the longest shortest path Science, 
%https://www.science.org/doi/full/10.1126/science.aau9735
%Distinction homophily, chicken or egg of

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle.png}
            \caption{Circle Graph}
            \label{fig:circle}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle-plus-triangle.png}
            \caption{Circle Graph with Closure}
            \label{fig:circleplus}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/circle-plus-long.png}
            \caption{Circle Graph with Long Tie}
            \label{fig:circleplus}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/clique.png}
            \caption{Clique Graph}
            \label{fig:clique}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/kite.png}
            \caption{Kite Graph}
            \label{fig:kite}
    \end{subfigure}
    \caption{Distinction Centrality in toy graphs. Nodes with the highest distinction are shown in red, nodes with positive distinction are shown in blue, nodes with negative distinction are shown in tan, and nodes with distinction close to zero are shown in purple.}
    \label{fig:toys}
\end{figure}

\subsubsection{Long Ties and Hierarchy}
This section highlights how the hierarchy is really formed out of the redundant connections of the subgraphs. While a star graph clearly creates a hierarchy, when additional weight is added to the broker's alters, the hierarchy can be flipped, especially when they form denser subgraphs. This is in keeping with the observed structural holes above.

@@
